Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-09-29 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
discussion

@@

While you are reviewing the scene you meet up with the fire captain and begin asking him about the fire.  The captain tells you that the doors and windows were closed upon arrival.  When the front door was opened heavy smoke

>>>>

Comparison of Fire Patterns and Flame Heights

Each of the two sets of experiments, the fire source characterization and the fire pattern, had uncertainties associated with them.  In the case of the fire source characterization, there were uncertainties associated with the fuel composition itself, the measurement of the heat release rate, temperature, heat flux and flame height.  The fire pattern experiments had additional uncertainties which included: the composition and thickness of the gypsum wallboard, moisture content and paint thickness.  In both sets of experiments, efforts were made to limit the uncertainties in the fuels, the measurement techniques and the gypsum wallboard. In this laboratory based study, the key question is – for a given fire pattern, is there a relationship that can be demonstrated to determine the source fire?

At this stage of the research, the answer to the above question is yes and no.  Figure 16 shows the comparisons of the average maximum fire pattern heights with those of the mean flame height based on the analysis of the video images.  The associated 95~\% confidence limits are also shown for each height.

The comparison for the fire pattern heights and flame heights from the natural gas fueled fires are very similar.  This is expected in part because the natural gas fires are “steady state”.  The “steady” nature of the fire aids in the repeatability or the reduction of variability in the experiments.  For example the energy transfer to the gypsum wall board is also fairly “steady” for the duration of the experiment.  Based on the average total heat released, approximately 23 MJ, more energy was transferred to the wall board when compared to the other two fuels.  This energy is needed to “dry” out the gypsum wallboard both of water absorbed from the environment and the chemically bound water, to allow the paper cover and paint to burn.

The gasoline fueled fires also had similar fire pattern heights and flame heights; however, the similarity was not as strong as the natural gas fires.  The transient nature of the fire played a strong role here, there was less energy transferred to the gypsum wall board and the exposure time for portions of the wall board more distant from the fuel surface was less.  Another issue that comes into the discussion is the length of time, 30 s, that was used for flame height analysis.  This period of time was used to ensure that the flame height at the time of peak heat release rate was included in the analysis.  However, this ultimately may have had the effect of reducing the mean flame height.  The mean flame height based on the photo analysis was 0.84 m.  The photo analysis was also performed over a 30 s period bounding the peak heat release rate but the number of images analyzed was about one sixth of the images analyzed with the video technique.
It should also be noted that the range of the flame heights increased, especially on the lower bound, with the video analysis compared to the photo analysis.

The polyurethane foam fueled fire heights and the associated fire pattern heights had the least similarity when compared to the other two fuels.  This follows with the other findings about the polyurethane fueled fires in this study.  The most prominent being the complications added by the fuel being solid and the uncertainties of the flame spread and localized burning rates, leading to poor repeatability of the fire growth and the heat release rate relative to the natural gas and the gasoline fueled fires.

Mean flame height has been identified by Heskestad and others as having a strong relation to a given heat release rate from fires with a horizontal fuel surface~\cite{Heskestad:SFPE, Beyler:1986}.  However, the ability to relate the mean flame height to the fire pattern height, based on the limited tests here is best suited to a “steady state” case.  In cases where transient fire growth must be considered, flame height alone may not be enough to determine the heat release rate of the fire without further information to enable an energy balance calculation on the gypsum wall board or other target fuel even for “simple” pre-flashover fire patterns.

@@

While you are reviewing the scene you meet up with the fire captain and begin asking him about the fire.  The captain tells you that the doors and windows were closed upon arrival.  When the front door was opened heavy smoke and flame 
flow path
inconsistant fire pattern near the right side of the sofa.  It appears that pattern was generated pre-flashover.  The pattern height is on the order of   calc HRR and FH for cushion.
